---
title: "R downloads"
author: "Anna Reynolds"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  rmarkdown::html_document:
    code_folding: show
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This weeks #TidyTuesday dataset contains data on R downloads over a single year from RStudio CRAN mirror between October 20, 2017 and October 20, 2018. Source: cran-logs.rstudio.com. The data contains date, time, size, version, operating system and country variables.

## Load data
```{r , warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(knitr)
library(here)
library(hrbrthemes)
library(gghighlight)
library(scales)
library(ggthemr)
library(waffle)
#Set theme as Dust using ggthemr
ggthemr(palette = "dust", layout = "clear", set_theme = TRUE)
#Import data from website
#Rdown <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-10-30/r_downloads_year.csv")
Rdown <- read_csv(here("R_downloads.csv"))
kable(head(Rdown))
```

```{r output tidy data, echo=FALSE}
#Lets output the tidy dataset for future use.
write.csv(Rdown, file = "R_downloads.csv")
```

## Analysis/visualization
### Which countries had the most R downloads?

```{r bar chart, fig.width=10, fig.height=8}
#Limit to top 20 countries
Rdown_20  <- Rdown %>% group_by(country) %>% summarise(country_n = n()) %>% top_n(country_n,n=20) %>% filter(country != "NA")
bar <- ggplot(Rdown_20, aes(reorder(country, country_n), y = country_n)) + geom_bar(stat = "identity") + coord_flip() +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = "Number of downloads",
  title = "Top 20 countries downloading R between October 20, 2017 and October 20, 2018",
  caption = "Source: cran-logs.rstudio.com")
bar
```


### Which operating system was most commonly used?
```{r waffle chart}
#Summarise downloads by operating system
os  <- Rdown %>% group_by(os) %>% summarise(os_n = n()) %>% mutate(percent = (os_n/sum(os_n))*100)
os_percent <- c(`OSX` = 21, `Windows` = 74, `Src` = 5)

#Make a waffle chart
waffle <- waffle(os_percent, rows=8, colors = c("#969696", "#1879bf", "#db735c"), legend_pos = "bottom", xlab = NULL, title = "R downloads (%) by operating system\nOctober 20, 2017 and October 20, 2018")
waffle
```

### Which version of R was most popular?
```{r bar2, fig.width=10, fig.height=6}
Rversion  <- Rdown %>% group_by(version) %>% summarise(version_n = n()) %>% top_n(version_n,n=10)
bar2 <-
  ggplot(Rversion, aes(x = reorder(version, version_n), y = version_n)) + geom_bar(stat = "identity") + coord_flip() +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = "Number of downloads",
  title = "Top 10 R versions downloaded between October 20, 2017 and October 20, 2018",
  caption = "Source: cran-logs.rstudio.com"
  )
  bar2
```

### When were the most common and least common dates for downloads?
```{r heat map, warning=FALSE, message=FALSE}
#Use lubridate to extract month and day?
library(lubridate)
Rdown2 <- Rdown %>% mutate(mth = month(date)) %>% mutate(day = day(date))
Rdown_date <- Rdown2 %>% group_by(mth, day) %>% summarise(n()) %>% rename(count = `n()`)
hm <- ggplot(data = Rdown_date, aes(x = mth, y = day)) +
  geom_tile(aes(fill = count))
hm
```

```{r line charts, warning=FALSE, message=FALSE}
library(plotly)
Rdown_date2 <- Rdown2 %>% group_by(date) %>% summarise(n()) %>% rename(count = `n()`)
line <- ggplot(Rdown_date2, aes(date, count)) + geom_line() + geom_smooth(se=FALSE, color = "steel blue") +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = " ",
  title = "R downloads between October 20, 2017 and October 20, 2018",
  caption = "Source: cran-logs.rstudio.com"
  )
line
#Make an interactive version of the line chart
ggplotly(line)
```


## Results
Between October 2017 and October 2018:

+ The most number of downloads originated from the United States
+ Windows was the most common operating system for downloading R
+ R versions 3.5.1 and 3.4.3 were the most common versions downloaded
+ September and October were the months that had the most downloads


## Discussion/conclusions
Downloads of R increased substantially in August 2018. Version 3.5.1 was released on 2 July 2018.

## Package citations

  Carson Sievert, Chris Parmer, Toby Hocking, Scott Chamberlain, Karthik Ram,
  Marianne Corvellec and Pedro Despouy (2017). plotly: Create Interactive Web
  Graphics via 'plotly.js'. R package version 4.7.1.
  https://CRAN.R-project.org/package=plotly
  
  Bob Rudis (2018). hrbrthemes: Additional Themes, Theme Components and
  Utilities for 'ggplot2'. R package version 0.5.0.
  https://CRAN.R-project.org/package=hrbrthemes
  
  Hadley Wickham (2017). tidyverse: Easily Install and Load the
  'Tidyverse'. R package version 1.2.1.
  https://CRAN.R-project.org/package=tidyverse
  
  Hadley Wickham (2018). stringr: Simple, Consistent Wrappers for Common
  String Operations. R package version 1.3.0.
  https://CRAN.R-project.org/package=stringr
  
  Yihui Xie (2018). knitr: A General-Purpose Package for Dynamic Report
  Generation in R. R package version 1.20.
  
  Bob Rudis and Dave Gandy (2017). waffle: Create Waffle Chart Visualizations in
  R. R package version 0.7.0. https://CRAN.R-project.org/package=waffle
  
  Ciaran Tobin (NA). ggthemr: Themes for ggplot2. R package version 1.1.0.
  
```{r citation, echo=FALSE, eval=FALSE}
citation("plotly")
citation("waffle")
citation("hrbrthemes")
citation("tidyverse")
citation("stringr")
citation("knitr")
citation("ggthemr")
```

