---
title: "Fast food calories"
author: "Anna Reynolds"
date: "6 September 2018"
output: 
  rmarkdown::html_document:
    code_folding: show
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(hrbrthemes)
```

## Fast food calories in the US
This weeks #TidyTuesday data set is on calories in fast food meals in restaurant chains in the US.

### Import the data
```{r import data, warning=FALSE, message=FALSE}
fastfood <- read_tsv("fastfood_calories.csv")
head(fastfood)

```

### Plot the calories for each fast food chain

#### Which chain has the widest distribution of products in terms of calories?

```{r beeswarm plot, fig.width=12}
library(beeswarm)
beeswarm(calories ~ restaurant, data=fastfood, col=2:9, pch=19, method="swarm", cex=1, horizontal=F, xlab="Data source: fastfoodnutrition.org", ylab="Calories", main="Distribution of calories in items, by restaurant chain", bty="n")
```

McDonalds had the largest distribution of calories with 4 items containing over 1500 calories. Most of the top 10 items at McDonalds by calories are chicken meals (e.g. tenders or nuggets).

```{r bar top10, fig.width=12}
mcdonalds <- filter(fastfood, restaurant == "Mcdonalds")
mcds_10 <- mcdonalds  %>% top_n(calories,n=10)
bar1 <- ggplot(mcds_10, aes(x = reorder(item, calories), y = calories))+ geom_bar(stat = "identity") + coord_flip() + theme_ipsum_rc() + labs(x=" ", y="Calories",
       title="Top 10 items by calories, McDonalds, USA",
       caption="Source: fastfoodnutrition.org")
bar1
```

### What about sodium?
```{r beeswarm plot2, fig.width=12}
beeswarm(sodium ~ restaurant, data=fastfood, col=2:9, pch=19, method="swarm", cex=1, horizontal=F, xlab="Data source: fastfoodnutrition.org", ylab="Sodium (mg)", main="Distribution of sodium in items, by restaurant chain", bty="n") 

```

McDonalds items also have a wide distribution of sodium content. There are a number of items above the recommended daily sodium intake of 2,300 mg per day https://www.cdc.gov/salt/pdfs/sodium_dietary_guidelines.pdf. The items that have high calories also have high sodium content.

```{r bar sodium, fig.width=12}
mcds_10_sodium <- mcdonalds  %>% top_n(sodium,n=10)
bar2 <- ggplot(mcds_10_sodium, aes(x = reorder(item, sodium), y = sodium))+ geom_bar(stat = "identity") + coord_flip() + theme_ipsum_rc() + labs(x=" ", y="Sodium (mg)",
       title="Top 10 items by sodium content, McDonalds, USA",
       caption="Source: fastfoodnutrition.org")
bar2 + geom_hline ( yintercept = 2300, size=1 ) + annotate ( "text" , x = 1 , y = 3100 , label = "Recommended daily intake")
```

### Is there a relationship between calories and sodium content?
```{r scatter, fig.width=12, fig.height=12}
ggplot ( fastfood , aes ( x = calories , y = sodium)) + geom_point () + theme_ipsum_rc() + labs(x="Calories ", y="Sodium (mg)",
       title="Calories vs sodium content, fast food companies, USA",
       caption="Source: fastfoodnutrition.org") + facet_wrap ( ~ restaurant )
```

Taco bell has the most items with lower calories and sodium.
