---
title: "Global mortality"
author: "Anna Reynolds"
date: "17 April 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)
library(hrbrthemes)
library(viridis)
library(RColorBrewer)
library(plotly)
library(ggridges)
```

## Global mortality 1990 to 2016

This weeks Tidy Tuesday data set contains data on global causes of death. It will need to be tidied using the tidyverse package before it can be graphed.

### Import the data

```{r import data set}
world_mort <- readxl::read_excel('global_mortality.xlsx') 
head(world_mort)
```

## Tidy the data

The cause columns contain (%) which will need to be removed before the gather function from dplyr is used to convert the data set from wide to long.

```{r tidy the data}
#Remove % from column names
world_mort2 <- rename_at(world_mort,vars(contains("%")),funs(str_remove_all(.,"[[:punct:]]"))) 
world_mort_tidy <- gather(world_mort2, `Cardiovascular diseases `:`Terrorism `, key="cause", value = "percent")
```

Lets recreate the horozontal bar chart for global deaths. Cardiovascular diseases are responsible for the highest share of deaths world wide in 2016.

```{r bar chart world, fig.height=8}
world <- filter(world_mort_tidy, country == "World" & year == 2016) 

ggplot(world, aes(x = reorder(cause, percent), y = percent, fill = percent)) + geom_bar(stat = "identity") + coord_flip() + scale_fill_viridis(direction=-1) + theme_minimal() + labs(x="Cause", y="Percent",
       title="Share of deaths by cause, World, 2016",
       caption="Source: ourworldindata.org; IHME GBD Study")
```

How have deaths from Cardiovacular diseases changed over time?

```{r line chart, fig.height=8}
#Restrict the data to Cardiovascular diseases and SDI
cvd_sdi <- world_mort_tidy %>% filter(str_detect(country, "SDI") & cause == "Cardiovascular diseases ")

ggplot(cvd_sdi, aes(x = factor(year), y = percent, group = country, colour = country)) + geom_line(size = 1.25) + theme_minimal() + labs(x="Year", y="Percent",
       title="Cardiovascular disease share of deaths, by income groups, 1990-2016",
       caption="Source: ourworldindata.org; IHME GBD Study")

```

Between 1990 and 2016, the share of deaths due to Cardiovascular diseases has declined in High SDI countries, been stable in High-middle SDI countries and increased in Middle, Low-middle and Low SDI countries.


```{r output tidy data, echo=FALSE}
#Lets output the tidy dataset for future use.
write.csv(world_mort_tidy, file = "global_mortality_tidy.csv")
```