---
title: "Global life expectancy"
author: "Anna Reynolds"
date: "9 July 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(hrbrthemes)
library(ggthemes)
library(RColorBrewer)
display.brewer.all(colorblindFriendly=TRUE)
library(ggpomological)
```


## TidyTuesday - Global life expectancy

This weeks dataset contains data on global life expectancy from Our World in Data <https://ourworldindata.org/life-expectancy> Max Roser (2018).  The date range is 1950 to 2015.

```{r import data}
le_global <- read_csv('week14_global_life_expectancy.csv', col_names = TRUE)
head(le_global)
```

### Lets look at the high level picture by global region

Restrict the data to Africa, Asia, Oceania, the Americas, Europe, and the World.

```{r high level}
le_high <- le_global %>% filter(country == 'Asia' | country == 'Africa' | country == 'Oceania' | country == 'Europe' | country == 'World' | country == 'Northern America' | country == 'South America')
```

```{r line chart, fig.width=8}
cb_palette <- c ( "#999999" , "#E69F00" , "#56B4E9" , "#009E73" , "#000000" , "#0072B2" , "#D55E00" , "#CC79A7" )

ggplot(le_high, aes(x = factor(year), y = life_expectancy, group = country, color = country)) + geom_line(size = 1.25) + theme_economist_white() + scale_color_manual ( values = cb_palette ) + labs(x="Year", y="Life expectancy",
       title="Global life expectancy, 1950 to 2015",
       caption="Source: https://ourworldindata.org/life-expectancy", color=" ") + scale_x_discrete ( breaks = c ( 1950 , 1970, 1990 , 2010)) 
```

Life expectancy has increased over time in all regions, however Africa still lags well behind the rest of the world.  

```{r save plot as png, echo=FALSE}
ggsave ("life_exp_region.png", width=8, height=6, dpi=300)
```
### ggpomological version
```{r ggpomological, fig.width=8}
ggplot(le_high, aes(x = factor(year), y = life_expectancy, group = country, color = country)) + geom_line(size = 1.25) + theme_pomological_fancy() + labs(x="Year", y="Life expectancy",
       title="Global life expectancy, 1950 to 2015",
       caption="Source: https://ourworldindata.org/life-expectancy", color=" ") + scale_x_discrete ( breaks = c ( 1950 , 1970, 1990 , 2010)) 
```

### Lets have a look at the impact of war and conflict on life expectancy.
Restrict data to countries that have experienced conflict/war during the time period.
```{r war line chart, fig.width=8}
le_war <- le_global %>% filter(country == 'Cambodia' & year >1945 | country == 'Vietnam' | country == 'Rwanda' | country == 'Syria' | country == 'World' | country == 'Iraq' | country == 'Ethiopia')

ggplot(le_war, aes(x = factor(year), y = life_expectancy, group = country, color = country)) + geom_line(size = 1.25) + theme_economist_white() + scale_color_brewer ( palette = "Set2" ) + labs(x="Year", y="Life expectancy",
       title="Global life expectancy, 1950 to 2015",
       subtitle="What is the effect of war/conflict?",
       caption="Source: https://ourworldindata.org/life-expectancy", color=" ") + scale_x_discrete ( breaks = c ( 1950 , 1970, 1990 , 2010, 2015)) 

```

Life expectancy has been most affected by the Khmer Rouge regime in Cambodia in the late 1970's and the civil war in Rwanda in the mid 1990's.
expectancy.
```{r save as png, echo=FALSE}
ggsave ("life_exp_war.png", width=8, height=6, dpi=300)
```