---
title: "US census"
author: "Anna Reynolds"
date: "5 May 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)
library(hrbrthemes)
library(viridis)
library(RColorBrewer)
library(plotly)
library(ggridges)
library(scales)
```

## US census data - Tidy Tuesday

This week the data for Tidy Tuesday comes from the US Census. It contains data by US county and there are a number of continuous variables such as unemployment rate, poverty rate and population data. This week I might try a scatterplot or two.

```{r import data}
census_data <- read.csv('us_census.csv', stringsAsFactors = FALSE, header=TRUE, sep=",", na.strings="M")
head(census_data)
tail(census_data)
```

<details>
<summary>View a summary of the US census data variables</summary>
```{r summary, echo=FALSE}
summary(census_data)
```
</details>

## Plot of Income vs Poverty 
```{r scatterplot, fig.height=8}
#Make a variable with County and State combined
census_data2 <- unite(census_data, col='CountyState', County, State, sep=", ", remove=FALSE)

sp <- ggplot (census_data2, aes ( x = IncomePerCap , y = Poverty, color = Unemployment, label=CountyState)) + geom_point () + theme_minimal() + scale_colour_gradientn ( colours = c ("yellow", "orange", "darkred")) + labs(x="Income per capita ($US)", y="Poverty (%)", title="Income vs poverty, US counties", subtitle="County-level American Community Survey (5-year estimates) 2015", caption="Data source: census.gov")
sp
```

## An interactive version
```{r interactive scatterplot, message=FALSE, warning=FALSE, fig.width=8}
ggplotly(sp, tooltip = c("label", "x", "y", "color"))
```

Most of the counties that have low income per capita and high poverty and unemployment rates are in Puerto Rico.

## Distribution of income per capita by state
```{r ridge plot, fig.height=8, warning=FALSE, message=FALSE}

ggplot(census_data2, aes(x = IncomePerCap, y = State)) + geom_density_ridges( color="#FFFFFF", fill="#56B4E9") + theme_minimal() + labs(x="Income per capita ($US)", title = "Income per capita distribution, USA, 2015",
    caption = "Data source: census.gov")
```

```{r box plot, fig.height=8, warning=FALSE, message=FALSE}

ggplot(census_data2, aes(x = fct_reorder(State, IncomePerCap, fun = median), y = IncomePerCap)) + geom_boxplot(fill="#56B4E9") + theme_minimal() + labs(x = "State", y="Income per capita ($US)", title = "Income per capita distribution, USA, 2015",
    caption = "Data source: census.gov") + coord_flip() 
```