---
title: "US College majors"
author: "Anna Reynolds"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  rmarkdown::html_document:
    code_folding: show
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This #TidyTuesday data set contains the data behind the story [The Economic Guide To Picking A College Major](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/){target="_blank"} by Fivethirtyeight. All data is from American Community Survey 2010-2012 Public Use Microdata Series.

## Load data
```{r , warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(knitr)
library(hrbrthemes)
library(gghighlight)
library(scales)
theme_set(theme_ipsum_rc())
college_majors <- read_csv("college_majors.csv") %>% mutate(Major = str_to_title(Major)) %>% mutate(Women_percent=ShareWomen*100)
kable(head(college_majors))
```

## Analysis/visualization
### What is the most popular major?

```{r bar chart, fig.width=8, fig.height=10}
top_20 <- college_majors  %>% top_n(Total,n=20)
bar1 <-
  ggplot(top_20, aes(x = reorder(Major, Total), y = Total)) + geom_bar(stat = "identity") + coord_flip() +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = "Number of people",
  title = "Top 20 popular majors 2010-2012, USA",
  caption = "Source: American Community Survey 2010-2012 Public Use Microdata Series"
  )
  bar1
```

### Which majors have the highest and lowest proportions of women?
#### Highest proportions of women.
```{r bar chart 2, fig.width=8, fig.height=6}
top_20_women <- college_majors  %>% top_n(ShareWomen,n=20) %>% mutate(Women_percent=ShareWomen*100)

bar2 <-
  ggplot(top_20_women,
  aes(
  x = reorder(Major, Women_percent),
  y = Women_percent,
  fill = Major_category
  )) + geom_bar(stat = "identity") + coord_flip() + labs(
  x = " ",
  y = "Percent",
  title = "Top 20 majors with highest proportion\nof women, 2010-2012, USA",
  caption = "Source: American Community Survey 2010-2012 Public Use Microdata Series"
  ) + scale_fill_discrete(name = "Major category") +
    scale_fill_manual(values=c("#FB9A99", "#6A3D9A", "#FFFF99", "#B15928", "#CAB2D6"))
  bar2
```

#### Lowest proportions of women.
```{r bar chart 3, fig.width=8, fig.height=6}
bottom_20_women <- college_majors  %>% top_n(ShareWomen,n=-20) %>% mutate(Women_percent=ShareWomen*100)
bar3 <-
  ggplot(bottom_20_women,
  aes(
  x = reorder(Major, Women_percent),
  y = Women_percent,
  fill = Major_category
  )) + geom_bar(stat = "identity") + coord_flip() + labs(
  x = " ",
  y = "Percent",
  title = "Bottom 20 majors with lowest proportion\nof women, 2010-2012, USA",
  caption = "Source: American Community Survey 2010-2012 Public Use Microdata Series"
  ) + scale_fill_discrete(name = "Major category") +
    scale_fill_brewer(palette="Paired")
  bar3
```

### Which majors have the highest incomes?
```{r bar chart 4, fig.width=10, fig.height=8}
top_20_inc <- college_majors  %>% top_n(Median,n=20)
bar4 <-
  ggplot(top_20_inc, aes(x = reorder(Major, Median), y = Median)) + geom_bar(stat = "identity") + coord_flip() +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = "Median Income ($)",
  title = "Top 20 majors by median income 2010-2012, USA",
  caption = "Source: American Community Survey 2010-2012 Public Use Microdata Series"
  )
  bar4
```

### Rank by proportion of women
```{r scatterplot, fig.width=10, fig.height=8, warning=FALSE, message=FALSE}
scatter <-
  ggplot(college_majors,
  aes(x =  Women_percent, y = Rank, color = Major_category)) + geom_point() + labs(
  x = "Percent",
  y = "Rank",
  title = "College majors by proportion of women and income (rank), \n2010-2012, USA",
  caption = "Source: American Community Survey 2010-2012 Public Use Microdata Series"
  ) + scale_color_discrete(name = "Major category") +
  scale_y_reverse()
scatter
```

```{r scatterplot2, warning=FALSE}
scatter2 <-
  ggplot(college_majors, aes(x =  Women_percent, y = Rank)) + geom_point(size =
  2) + gghighlight(Rank < 11) + labs(
  x = "Percent",
  y = "Rank",
  title = "College majors by proportion of women and income (rank), \n2010-2012, USA",
  subtitle = "Top 10 majors by salary",
  caption = "Source: American Community Survey 2010-2012 Public Use Microdata Series"
  ) + annotate(
  "text",
  label = "Astronomy and Astrophysics",
  x = 71,
  y = 8,
  size = 3.5
  ) + scale_y_reverse()
scatter2
```

## Conclusions
+ Psychology was the most popular major followed by Business Management and administration
+ Education and Health majors had the highest proportion of women
+ Engineering majors had the lowest proportion of women but the highest salaries
+ Petroleum engineering had the highest median income
+ The higher the income rank of a major, the lower the proportion of women
+ Astronomy and astrophysics was an outlier, with more than 50% women and ranked number 8 for income

## Package citations
  Hiroaki Yutani (2018). gghighlight: Highlight Lines and Points in
  'ggplot2'. R package version 0.1.0.
  https://CRAN.R-project.org/package=gghighlight
  
  Bob Rudis (2018). hrbrthemes: Additional Themes, Theme Components and
  Utilities for 'ggplot2'. R package version 0.5.0.
  https://CRAN.R-project.org/package=hrbrthemes
  
  Hadley Wickham (2017). tidyverse: Easily Install and Load the
  'Tidyverse'. R package version 1.2.1.
  https://CRAN.R-project.org/package=tidyverse
  
  Hadley Wickham (2018). stringr: Simple, Consistent Wrappers for Common
  String Operations. R package version 1.3.0.
  https://CRAN.R-project.org/package=stringr
  
  Yihui Xie (2018). knitr: A General-Purpose Package for Dynamic Report
  Generation in R. R package version 1.20.
  
```{r citation, echo=FALSE, eval=FALSE}
citation("gghighlight")
citation("hrbrthemes")
citation("tidyverse")
citation("stringr")
citation("knitr")
```

