---
title: "Dallas animal shelter"
author: "Anna Reynolds"
date: "10 August 2018"
output: 
  rmarkdown::html_document:
    code_folding: show
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dallas animal shelter data

This weeks #TidyTuesday data set is on animal shelter data in Dallas in the US. There is one Excel spreadsheet with data from the 2016-17 financial year.

### Import the data

```{r import data, warning=FALSE, message=FALSE}
library(tidyverse)
library(paletteer)
library(hrbrthemes)
library(ggalluvial)

animals <- readxl::read_excel(sheet = "simple", 'week18_dallas_animals.xlsx')
head(animals)

```

### Which animal type was most often taken in by the shelter?
```{r type, fig.height=8}
type <- ggplot(animals, aes(x = animal_type)) + geom_bar(stat = "count") + coord_flip() + theme_ipsum_rc() + labs(x="Type of animal", y="Number of animals",
       title="Dallas animal shelter data, USA, 2016-17",
       caption="Source: Dallas OpenData")
type
```

Dogs were the most common animal type taken in by the shelter.

### What was the most common breed of dog at the shelter?
```{r bar dogs, fig.height=8, fig.wdith=12, message=FALSE, warning=FALSE}
dogs <- animals %>% filter(animal_type=="DOG") %>% count(animal_breed) %>% top_n(n=20)
dog_breeds <- ggplot(dogs, aes(x = reorder(animal_breed, n), y = n)) + geom_bar(stat = "identity") + coord_flip() + theme_ipsum_rc() + labs(x="Dog breed", y="Number of animals",
       title="Dallas animal shelter data - top 20 dog breeds, USA, 2016-17",
       caption="Source: Dallas OpenData")
dog_breeds
```

Pitt Bulls, Short haired Chihuahua, German Sheperds and Labrador Retrievers were the most common breeds.

### When were dogs most often taken in?
```{r line dogs, fig.height=8, fig.wdith=12, message=FALSE, warning=FALSE}
dogs_only <- animals %>% filter(animal_type=="DOG") %>% count(intake_date)
date <- ggplot(dogs_only, aes(x = intake_date, y = n, group=1)) + geom_line() + stat_smooth ( method = loess ) + theme_ipsum_rc() + labs(x="Intake date", y="Number of animals",
       title="Dallas animal shelter data - intake date for dogs, USA, 2016-17",
       caption="Source: Dallas OpenData")  
date
```

Intake of dogs peaked in the summer months.

```{r month bar, fig.height=8, fig.wdith=12, message=FALSE, warning=FALSE}
dog_month <- animals %>% filter(animal_type=="DOG") %>% count(month) %>% top_n(n=5)
month <- ggplot(dog_month, aes(x = reorder(month, n), y = n)) + geom_bar(stat = "identity") + coord_flip() + theme_ipsum_rc() + labs(x="Month", y="Number of animals",
       title="Dallas animal shelter data - month of intake, USA, 2016-17",
       caption="Source: Dallas OpenData")
month
```

June was the most common month of intake for dogs into the shelter.

### What are the outcomes for dogs?
```{r sankey, fig.height=8, fig.wdith=16, message=FALSE, warning=FALSE}
dogs_sum <- animals %>% filter(animal_type == "DOG") %>% filter(intake_type %in% c("STRAY", "OWNER SURRENDER", "CONFISCATED")) %>% mutate(outcome_type = fct_lump(outcome_type, n = 4)) %>% group_by(intake_type, outcome_type) %>% summarise(n = n())

pal <- paletteer_d(wesanderson, Moonrise3)

sankey <-
  ggplot(dogs_sum, aes(y = n, axis1 = intake_type, axis2 = outcome_type)) + geom_alluvium(aes(fill = outcome_type)) + geom_stratum(fill = "lightgrey", color = "white", size = 1) + geom_text(stat = "stratum", label.strata =
  T) + theme_minimal() + labs(
  x = " ",
  y = " ",
  title = "Dallas animal shelter data - outcomes for dogs, USA, 2016-17",
  caption = "Source: Dallas OpenData",
  fill = "Outcome"
  ) + scale_fill_manual(values = pal) + theme (panel.grid.major = element_blank (),
  panel.grid.minor = element_blank ())
  sankey

```
```{r colorblindr}
#Check if sankey is colour blind friendly
cvd_grid(sankey)
```

