---
title: "US airports"
author: "Anna Reynolds"
date: "22 September 2018"
output: 
  rmarkdown::html_document:
    code_folding: show
    theme: sandstone
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(hrbrthemes)
library(viridis)
library(RColorBrewer)
library(geofacet)
library(cartography)
library(sf)
library(knitr)
```

## US airports - Tidy Tuesday
This weeks #TidyTuesday dataset contains information about US airports from 2013-2017. More information about the data can be found here: https://www.faa.gov/airports/planning_capacity/passenger_allcargo_stats/categories/

### Import the data
```{r import data, message=FALSE}
#Import .csv file
airports <- read.csv('us_airports.csv')
kable(head(airports))
#Filter to 2016 and remove small US territories (e.g. Puerto Rico etc)
airports_16 <- airports %>% filter(year==2016 & (state != "GU" & state != "PR" & state != "VI" & state != "AS" & state != "MP"))
 
#Import the US states shapefile
us_states <- st_read("C:/Users/anna/Documents/Data_visualisation/Shapefiles/cb_2017_us_state_5m/cb_2017_us_state_5m.shp")
```

The data for 2017 may be incomplete so we will focus on 2016.

### In 2016 which states had the most airports?
```{r most airports}
number_airports <- airports_16 %>% group_by(state) %>% summarise(n=n()) %>% arrange(desc(n))
kable(head(number_airports))
```

Alaska has the most number of airports with 258, compared with California with 75 then Texas with 63.
Why does Alaska have so many airports?
Lets have a look at the number of airports by type.
```{r types airports}
type_air <- airports_16 %>% group_by(state, airport_classification) %>% summarise(type=n())
```

```{r geofacet, fig.width=12, fig.height=8, message=FALSE}
ggplot(type_air, aes(airport_classification, type, group = state)) +
  geom_bar(stat = "identity") + coord_flip() +
  facet_geo(~ state, grid = "us_state_grid2", label = "name") +
    labs(title = " ",
    caption = "Source: www.faa.gov",
    x = " ",
    y = " ") +
  theme(strip.text.x = element_text(size = 6))
```

```{r cities lat long, message=FALSE, warning=FALSE}
#Merge the US states shapefile with the airports_16 data
states_shp <- st_read("C:/Users/anna/Documents/Data_visualisation/Shapefiles/cb_2017_us_state_5m/cb_2017_us_state_5m.shp") %>% rename(state = STUSPS)

#Import the us cities data containing lat/long
cities <- read.csv('us_cities.csv') %>% rename(state = state_id)

# Join the values to the map data using state
us_airports <- left_join(states_shp, airports_16)

#Join city lat/long data
us_airports2 <- left_join(us_airports, cities)

#Filter to California
cali <- us_airports2 %>% filter(state == 'CA')
plot(st_geometry(cali))

```

```{r cali map, warning=FALSE}
ggplot(cali) + geom_sf(fill="white") + geom_point(aes(x=lng, y=lat, size=passengers, color=hub_type)) + labs(x=NULL, y=NULL, 
                  title="Airports in California, 2016",
                  subtitle=NULL,
                  caption="Source: www.faa.gov") + theme_minimal() 
```
```{r us map, fig.width=12, fig.height=12}
ggplot(us_airports2) + geom_sf(fill="white") + geom_point(aes(x=lng, y=lat, size=passengers, color=hub_type)) + labs(x=NULL, y=NULL, 
                  title="Airports in the USA, 2016",
                  subtitle=NULL,
                  caption="Source: www.faa.gov") + theme_minimal() +  coord_sf(xlim=c(-170, -60), ylim=c(20, 80)) 
```

