---
title: "US wind farm locations"
author: "Anna Reynolds"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  rmarkdown::html_document:
    code_folding: show
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This weeks #TidyTuesday data set contains location and characteristic information about wind turbines in the USA. Some potential questions:

+ How do newer installations compare to older turbines?
+ Which states/regions have the most wind farms?
+ Spread of wind-turbines over time?
+ Where is the most missing data?

## Load data
I am trying to get better at using the here package to make the code even more reporoducible.
```{r load data, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(ggthemr)
library(knitr)
library(scales)
library(leaflet)
library(DataExplorer)
#Load data
us_wind <- read_csv(here("US_wind.csv"), na = c("missing",-9999,"n/a")) %>% rename(longitude = xlong, latitude = ylat)
kable(head(us_wind))
#Set theme
ggthemr(palette = "fresh", set_theme = TRUE)
#Create a report using DataExplorer
#create_report(us_wind)
kable(introduce(us_wind))
```

## Analysis/visualization
### Where is the most missing data?
```{r data explorer}
plot_intro(us_wind)
#Which variables have the most missing data?
plot_missing(us_wind)
```

### Remove duplicates
```{r}
#Number of projects
wind_dedup <- us_wind %>% distinct(t_fips, p_name, .keep_all = TRUE)
kable(introduce(wind_dedup))
```

### Which states have the most wind farms?
```{r bar chart}
state_20  <- wind_dedup %>% group_by(t_state) %>% summarise(t_state_n = n()) %>% top_n(t_state_n,n=20)
bar <- ggplot(state_20, aes(reorder(t_state, t_state_n), y = t_state_n)) + geom_bar(stat = "identity") + coord_flip() +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = "Number of wind farms",
  title = "Top 20 states with wind farms",
  caption = "Source: usgs.gov")
bar
```

### Where are the wind farms located?
Interactive map
```{r interactive map, warning=FALSE, message=FALSE}
#Filter out Guam
us_wind2 <- wind_dedup %>% filter(t_state != "GU")
leaflet(us_wind2) %>% addTiles() %>%
  addCircleMarkers(
    radius = 4,
    color = "blue",
    stroke = FALSE, fillOpacity = 0.5,
    popup = ~t_county
  )
```

Cluster map
```{r cluster map, warning=FALSE, message=FALSE}
leaflet(data = us_wind2) %>% 
  setView(-97.3606922, 32.7084083, zoom = 4) %>% 
  addTiles() %>% addMarkers(clusterOptions = markerClusterOptions())
```

Heat map - number of turbines per farm
```{r heat map, warning=FALSE, message=FALSE}
library(leaflet.extras)
leaflet(data = us_wind2) %>% 
  setView(-97.3606922, 32.7084083, zoom = 4) %>% 
  addTiles() %>%
  addHeatmap(
    lng = ~longitude, lat = ~latitude, intensity = ~p_tnum,
    blur = 20, max = 0.05, radius = 15
  )
```


### How have the farms changed over time?
```{r , warning=FALSE, message=FALSE}
capacity <- wind_dedup %>% group_by(p_year) %>% summarize(avg_capacity = mean(t_cap, na.rm=TRUE))
turbines <- wind_dedup %>% group_by(p_year) %>% summarize(avg_turbines = mean(p_tnum, na.rm = TRUE), range_turbines = IQR(p_tnum))
height <- wind_dedup %>% group_by(p_year) %>% summarize(avg_height = mean(t_ttlh, na.rm = TRUE))

line <- ggplot(turbines, aes(p_year, avg_turbines)) + geom_line() +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = " ",
  title = "US wind farms - average number of turbines, 1981 to 2018",
  caption = "Source: usgs.gov"
  )
line

lollipop <- ggplot(capacity, aes(x=p_year, y=avg_capacity)) + 
  geom_point(size=4) + 
  geom_segment(aes(x=p_year, 
                   xend=p_year, 
                   y=0, 
                   yend=avg_capacity)) + 
  labs(x = " ",
  y = " ", title="US wind farms - average turbine capacity (kW), 1981 to 2018", 
       caption="Source: usgs.gov")
lollipop

bar <- ggplot(height, aes(p_year, avg_height)) + geom_bar(stat="identity") +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = " ",
  title = "US wind farms - average height of turbines (metres), 1981 to 2018",
  caption = "Source: usgs.gov"
  )
bar

```

### Do taller turbines result in higher capacity?
```{r scattert, warning=FALSE, message=FALSE}
scatter <- ggplot(wind_dedup, aes(t_ttlh, t_cap)) + geom_point() +
  scale_y_continuous(labels = comma_format()) + labs(
  x = "Height",
  y = "Capacity (kW)",
  title = "US wind farms - turbine height vs capacity, 1981 to 2018",
  caption = "Source: usgs.gov"
  )
scatter
```

### How has the establishment of wind farms changed over time?
```{r line chart, warning=FALSE, message=FALSE}
wind_date <- wind_dedup %>% group_by(p_year) %>% summarise(n()) %>% rename(count = `n()`)
line3 <- ggplot(wind_date, aes(p_year, count)) + geom_line() +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = " ",
  title = "US wind farms, 1981 to 2018",
  caption = "Source: usgs.gov"
  )
line3
```

## Results

+ 6.7% of the data were missing and the t_img_date variable had the most missing data
+ Texas, Iowa and California have the most wind farms
+ The average number of turbines per project has declined over time
+ The average turbine capacity (kW) has increased over time but there is lots of missing data
+ The average height of the turbines has increased over time
+ The higher the turbine the greater the capacity
+ The number of wind farms has increased since around 2000, and reached a peak in 2012

## Package citations

  Boxuan Cui (2018). DataExplorer: Data Explorer. R package version 0.7.0.
  https://CRAN.R-project.org/package=DataExplorer
  
  Ciaran Tobin (NA). ggthemr: Themes for ggplot2. R package version 1.1.0.

  Hadley Wickham (2017). tidyverse: Easily Install and Load the
  'Tidyverse'. R package version 1.2.1.
  https://CRAN.R-project.org/package=tidyverse
  
  Hadley Wickham (2018). stringr: Simple, Consistent Wrappers for Common
  String Operations. R package version 1.3.0.
  https://CRAN.R-project.org/package=stringr
  
  Yihui Xie (2018). knitr: A General-Purpose Package for Dynamic Report
  Generation in R. R package version 1.20.
  
  Joe Cheng, Bhaskar Karambelkar and Yihui Xie (2018). leaflet: Create
  Interactive Web Maps with the JavaScript 'Leaflet' Library. R package version
  2.0.1. https://CRAN.R-project.org/package=leaflet
  
  Bhaskar Karambelkar and Barret Schloerke (2018). leaflet.extras: Extra
  Functionality for 'leaflet' Package. R package version 1.0.0.
  https://CRAN.R-project.org/package=leaflet.extras
  
```{r citation, echo=FALSE, eval=FALSE}
citation("DataExplorer")
citation("ggthemr")
citation("tidyverse")
citation("stringr")
citation("knitr")
citation("leaflet")
citation("leaflet.extras")
```

