---
title: "US Exercise"
author: "Anna Reynolds"
date: "22 July 2018"
output: 
  rmarkdown::html_document:
    code_folding: show
    theme: sandstone
---

## Exercise in the US

This weeks #TidyTuesday data set is from the CDC and contains data on the proprotion of people meeting the guidelines for exercise in each US state. 

```{r libs, warning=FALSE, message=FALSE}
library(tidyverse)
library(hrbrthemes)
library(ggthemes)
library(geofacet)
library(ggthemr)
```

## Importing the data
The data this week comes in an Excel spreadsheet with a worksheet called `source` and one called `tidy` and I will be importing the tidy worksheet.

```{r data import, warning=FALSE, message=FALSE}
state_exercise <- readxl::read_excel(sheet = "tidy", 'week16_exercise.xlsx', col_types = c("skip", "guess", "guess", "guess", "numeric"))
head(state_exercise)
```

### Filter the dataset to show state average by sex and work status
```{r bar chart, fig.height=6, message=FALSE, warning=FALSE}
#Filter data to state average
state_avg <-
  state_exercise %>% filter(state == "state_average" & sex != "both")
  ggthemr(palette = "fresh")
  bar <-
  ggplot(state_avg, aes(x = sex, y = exercise, fill = work_status)) + geom_bar(stat = "identity", position = "dodge") + labs(
  x = "Sex",
  y = "Per cent",
  title = "Proportion of people who meet exercise guidelines \n by work status, USA, 2010-15",
  caption = "Source: CDC"
  )
  bar
```

In general it looks like more men meet the guidelines than women, and for both sexes more people who work meet the guidelines than those who don't work.

```{r geofacet, fig.width=12, fig.height=8, warning=FALSE, message=FALSE}
state_exercise2 <-
  state_exercise %>% filter(sex != "both" & state != "state_average")
  ggplot(state_exercise2, aes(sex, exercise, fill = work_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_geo( ~ state, grid = "us_state_grid2", label = "name")  +
  labs(
  title = "Proportion of people who meet exercise guidelines \n by work status, USA, 2010-15",
  caption = "Source: CDC",
  x = "Sex",
  y = "%"
  ) +
  theme(strip.text.x = element_text(size = 6)) + coord_flip()
```