---
title: "Malaria worldwide"
author: "Anna Reynolds"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  rmarkdown::html_document:
    code_folding: show
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This weeks #TidyTuesday is on Malaria data from [Our World in Data](https://ourworldindata.org/malaria). I have chosen to concentrate on the data set containing Malaria deaths by age across the world and time. It contains data from 1990 to 2016 for countries worldwide.

## Load data
```{r load data, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(ggthemr)
library(knitr)
library(scales)
library(DataExplorer)
malaria <- read_csv(here("malaria_deaths_age.csv"))
kable(head(malaria))
#Set theme
ggthemr(palette = "light", set_theme = TRUE)
kable(introduce(malaria))
```

## Analysis/visualization

### Deaths by age for the world - children under 5 impacted the most by malaria
```{r line chart}
world <- malaria %>% filter(country == "World")
line <- ggplot(world, aes(year, deaths, group = age_group, color = age_group)) + geom_line(size = 1.5) +
  scale_y_continuous(labels = comma_format()) + labs(
  x = " ",
  y = " ",
  title = "Malaria deaths worldwide by age group, 1990 to 2016",
  caption = "Source: ourworldindata.org"
  ) + theme(legend.title = element_blank()) + 
  scale_color_discrete(breaks=c("Under 5", "5-14", "15-49", "50-69", "70 or older"))
line
```

Interactive line chart
```{r interactive line chart, message=FALSE}
library(plotly)
ggplotly(line)
```

## Results
+ Children aged under 5 have a much higher number of deaths than the other age groups
+ Deaths in children under 5 peaked in 2003 with around 750,000 but declined to around 510,000 in 2016.

## Discussion/conclusions

## Package citations

  Boxuan Cui (2018). DataExplorer: Data Explorer. R package version 0.7.0.
  https://CRAN.R-project.org/package=DataExplorer
  
  Ciaran Tobin (NA). ggthemr: Themes for ggplot2. R package version 1.1.0.

  Hadley Wickham (2017). tidyverse: Easily Install and Load the
  'Tidyverse'. R package version 1.2.1.
  https://CRAN.R-project.org/package=tidyverse
  
  Hadley Wickham (2018). stringr: Simple, Consistent Wrappers for Common
  String Operations. R package version 1.3.0.
  https://CRAN.R-project.org/package=stringr
  
  Yihui Xie (2018). knitr: A General-Purpose Package for Dynamic Report
  Generation in R. R package version 1.20.
  
  Carson Sievert, Chris Parmer, Toby Hocking, Scott Chamberlain, Karthik Ram,
  Marianne Corvellec and Pedro Despouy (2017). plotly: Create Interactive Web
  Graphics via 'plotly.js'. R package version 4.7.1.
  https://CRAN.R-project.org/package=plotly
  
```{r citation, echo=FALSE, eval=FALSE}
citation("DataExplorer")
citation("ggthemr")
citation("tidyverse")
citation("stringr")
citation("knitr")
citation("plotly")
```